Bringing the Power of eBPF to Open vSwitch

### 2018-10-22 paper.tex
afxdp experiment

@@

\subsection{Performance Evaluation}
% testbed

@@

Compared to xdpsock, ovs-vswitchd adds on top of it a UMEM memory pool,
and packets parsing, lookup and action execution.
Our initial prototype shows pretty terrible performance, only $0.5Mpps$ rxdrop.
We investigated and our first fix is to enable OVS's PMD-mode,
Poll-Mode-Driver, to the afxdp netdev.
In OVS, a non-pmd mode netdev does packet receiption by putting all the receiving
netdev's file descript together and ovs-vswitch polls them when there is an fd ready
to perform I/O.  As a result, the fd of the XSK is {\em shared} with other fds, and
we observed that the poll system call has pretty high overhead.

@@

For every stage of the optimization, we use Linux perf extensively, e.g., perf stat
and perf record/report.  With the above new design, we observe the new bottleneck
is at the UMEM memory pool we introduce to maintain the UMEM freelist, and the
packet metadata allocation, struct dp\_packet, for every receiving packet.
We introduce two major API, umem\_elem\_get(), which gets N free element from the
UMEM freelist, and umem\_elem\_put(), which places back the free UMEM buffer to
the freelist.
producer 
consumer
We implemented three data structures of UMEM memory pool as below
to compare their performance
\begin{itemize}
\item list\_head: Similar to Linux kernel's list\_head, each element contains it
UMEM addr, and a next pointer points to next free element. Elements are sparse in
memory. Consumer always takes the first element, head, and producer also place back
from the head of the list.
\item ptr\_ring: Similar to Linux kernel's ptr\_ring, an array of elements are
allocated in continuous memory region, and pointers (head and tail) point for
Consumer takes elements from the tail and when placing back, producer, puts
them into where head pointer points to.
\item ptr\_stack: Similar to ptr\_ring, but consumer and producer always use
the top few element, thus similar to push pop. Compared

\end{itemize}
% lock
% push/pop
% preallocate array
Since the UMEM

We then observe that the UMEM memory pool implementation ha

PMD
HugePage
UMEM element list
Batching
2-CPU operation
non-blocking send
non-block receive
pre-allocation dp\_packet
rxhash

\mycomment{
/* PMD: Poll modes drivers.  PMD accesses devices via polling to eliminate
 * the performance overhead of interrupt processing.  Therefore netdev can
 * not implement rx-wait for these devices.  dpif-netdev needs to poll
 * these device to check for recv buffer.  pmd-thread does polling for
 * devices assigned to itself.
 *
 * DPDK used PMD for accessing NIC.
 *}

>>>>

Table~\ref{compare} shows the difference between two projects:
\begin{table}
\centering
\scriptsize
\begin{tabular}{|c | c | c|}
\hline
 {\bf Projects} & {\bf OVS-eBPF } & {\bf OVS-AFXDP} \\ \hline\hline
%  Performance & $<$2Mpps & $>14$Mpps \\ \hline
  Performance & Low & High \\ \hline
  Driver Support & No & Yes \\ \hline
  Minimal Kernel & 4.18 & Unknown \\ \hline
  BPF Code Size & Large & Minimal \\ \hline
  Extensibility & Low & High \\ \hline
  \end{tabular}
\caption{\footnotesize
Comparison of the OVS-eBPF project and OVS-AFXDP project.
}
\label{compare}
\end{table}

@@

Define:
baseline rxdrop vs OVS-AFXDP rxdrop
baseline l2fwd vs OVS-AFXDO l2fwd
\subsection{Performance Evaluation}

\subsubsection{Testbed}

@@

Compared to xdpsock, ovs-vswitchd adds on top of it a UMEM memory pool,
and packets parsing, lookup and action execution.
Our initial prototype shows pretty terrible performance, only $0.5Mpps$ rxdrop.
We investigated and our first fix is to enable OVS's PMD-mode,
Poll-Mode-Driver, to the afxdp netdev.

\subsubsection{PMD netdev}
In OVS, a non-pmd mode netdev does packet receiption by putting all the receiving
netdev's file descript together and ovs-vswitch polls them when there is an fd ready
to perform I/O.  As a result, the fd of the XSK is {\em shared} with other fds, and
we observed that the poll system call has pretty high overhead.

@@

\subsubsection{Memory Management and Locks}
For every stage of the optimization, we use Linux perf extensively, e.g., perf stat
and perf record/report.  With the above new design, we observe the new bottleneck
is at the UMEM memory pool we introduce to maintain the UMEM freelist, and the
packet metadata allocation, struct dp\_packet, for every receiving packet.
We introduce two major API, umem\_elem\_get(), which gets N free element from the
UMEM freelist, and umem\_elem\_put(), which places back the free UMEM buffer to
the freelist.
producer 
consumer
We implemented three data structures of UMEM memory pool as below
to compare their performance
\begin{itemize}
\item list\_head: Similar to Linux kernel's list\_head, each element contains it
UMEM addr, and a next pointer points to next free element. Elements are sparse in
memory. Consumer always takes the first element, head, and producer also place back
from the head of the list.
\item ptr\_ring: Similar to Linux kernel's ptr\_ring, an array of elements are
allocated in continuous memory region, and pointers (head and tail) point for
Consumer takes elements from the tail and when placing back, producer, puts
them into where head pointer points to.
\item ptr\_stack: Similar to ptr\_ring, an array of elements are allocated
in continuous memory region, but both consumer and producer takes elements
from beginning of the array.
\end{itemize}

Although we allocate one UMEM per netdev, there might be multiple queues per
netdev sharing the same UMEM. As a result, the above three data structures
all require a mutex lock before accessing UMEM freelist. Linux perf reports
pthread\_mutex\_lock as one of the top CPU utilization function, and overall
shows around $2Mpps$ overhead for rxdrop experiments.
We change our design by 1) allocating per-queue UMEM region and 2)
allocating one pmd thread per queue. As a result, no lock is needed because
the each queue has only one thread and its own set of UMEM elements.

%\subsubsection{Metadat Allocation}
With the above change, Linux perf shows that the packets metadata allocation
takes a lot of CPU cycles, the dp\_packet\_init, dp\_packet\_set\_data.
So instead of allocating packet metadata at packet reception time, we
implement two data structures and compare their performance:
\begin{itemize}
\item Inlining in UMEM buffer: Since we already allocate 2K chunk for each
UMEM buffer, we reserve the first 256-byte in each UMEM element as struct dp\_packet
and initialize the dp\_packet as much as we can at allocation time.
So when packets arrive, each packet already has its packet metadata at a const
offset ahead.
\item Array on heap: This design simply allocate a continuous memory region
storing an array of packet metadata, and initialize them as much as we can.
\end{itemize}

%complete separation of the allocation of metadata structures from the allocation of packet data buffers.
% see mbuf dpdk https://doc.dpdk.org/guides/prog_guide/mbuf_lib.html
%Embed metadata within a single memory buffer the structure followed by a fixed size area for the packet data.
%Use separate memory buffers for the metadata structure and for the packet data.

With the above design change, we found that using the ptr\_stack with
array on heap get the best performance. Largely due to both data structures
have better spatial locality and more batching friendly.
For example, setting up 32 packet metadatas in an array definitely incurs
less cache misses than setting up in 32 elements in UMEM.
And getting 32 free UMEM elements from ptr\_stack has similar benefits.
With the above decision, the OVS rxdrop shows $17Mpps$, only $2Mpps$ less
than the baseline xdpsock rxdrop.

\subsubsection{Batching Send Syscall}
%574.790 M/sec  
We continued measured the performance of OVS-AFXDP l2fwd and.
observed only $4Mpps$, compared to $17Mpps$ baseline xdpsock.
Again, by analysis using perf, we found that the OVS pmd thread under rxdrop
has much fewer context switches compared to the l2fwd, indicating that
the PMD process spends much more time in kernel space than in userspace.
By using strace, we saw that the l2fwd experiement calls sendto system call
at very high frequeuce, because from Figure~\ref{afxdptx}, we design to check
the success of send (step 4) immediately after issusing send (step 3).
We change the design by calling send syscall (step 3) only when TX ring is
closed to full, e.g., when 3/4 ring elements have been used.
In specific case, instead of issuing send syscall for a batch of 32 packets
and making sure they are done, we only issue send when there are 512 outstanding
packets.  With this change, the OVS-AFXDP l2fwd experiment shows around $14Mpps$.

%rxhash
